Stability and accuracy of free surface time integration in viscous flows

### 2016-05-27 free-surface-paper.tex
Unnecessary sentence.
Thank reviewers.
Add note about timestepping in AMR test.

One downside of the quasi-implicit scheme is that it requires a modification of the system matrix.
The implementation of such a scheme results in a slightly 
asymmetric matrix, which can be more difficult to solve, requiring either
changes to the solver/preconditioner, or symmetrization of the stabilization term \citep{kaus2010stabilization}.
An alternative would be to construct a time-integration scheme that remains stable at larger timesteps.
The reason that forward Euler integration performs so badly with the decay of topography is that 
at larger timesteps it overshoots its equilibrium position. This overshoot causes it to lurch back 
to the \emph{other} side of equilibrium, overshooting even more.
We would like to construct an explicit scheme that accounts for following properties which we know the system has:
\begin{itemize}
\item In the absence of forcing, topography always decreases.
\item Relaxation of small amplitude topography takes the form of exponential decay.
\item The decay mode with the shortest relaxation time is the least stable.
\end{itemize}

@@

We have chosen a very small timestep for this test of 500 yr (note that this is significantly smaller than the smallest
relaxation time of $\sim$ 14 kyr), so most timestepping schemes should be stable and accurate, including forward Euler.
We choose this step for two reasons: (1) The primary purpose of this test is to investigate the savings
due to using adaptive mesh refinement with a free surface, and (2) \citet{crameri2012comparison}
found that a timestep of several hundred years was required to achieve full convergence.

@@

\section*{Acknowledgements}
This work was supported through National Science Foundation grant EAR-1246670
and by the Computational Infrastructure in Geodynamics (CIG), through National Science Foundation grant EAR-0949446. 
TH is partially supported by NSF grant DMS-1522191.
Thanks to Wolfgang Bangerth, Brent Delbridge, and Cedric Thieulot for useful discussions.
%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

>>>>

The implementation of such a scheme results in a slightly 
asymmetric matrix, which can be more difficult to solve, requiring either
changes to the solver/preconditioner, or symmetrization of the stabilization term \citep{kaus2010stabilization}.
An alternative would be to construct a time-integration scheme that remains stable at larger timesteps.
The reason that forward Euler integration performs so badly with the decay of topography is that 
at larger timesteps it overshoots its equilibrium position. This overshoot causes it to lurch back 
to the \emph{other} side of equilibrium, overshooting even more.
We would like to construct an explicit scheme that accounts for following properties which we know the system has:
\begin{itemize}
\item In the absence of forcing, topography always decreases.
\item Relaxation of small amplitude topography takes the form of exponential decay.
\item The decay mode with the shortest relaxation time is the least stable.
\end{itemize}

@@

We run simulations using both quasi-implicit and nonstandard finite difference schemes for the free surface.
We use a very small timestep for this test of 500 yr (note that this is significantly smaller than the smallest
relaxation time of $\sim$ 14 kyr), so most timestepping schemes should be stable and accurate, including forward Euler.
We choose this step for two reasons: (1) The primary purpose of this test is to investigate the savings
due to using adaptive mesh refinement with a free surface, and (2) \citet{crameri2012comparison}
found that a timestep of several hundred years was required to achieve full convergence.

@@

\section*{Acknowledgements}
This work was supported through National Science Foundation grant EAR-1246670
and by the Computational Infrastructure in Geodynamics (CIG), through National Science Foundation grant EAR-0949446. 
TH is partially supported by NSF grant DMS-1522191.
Thanks to Wolfgang Bangerth, Brent Delbridge, and Cedric Thieulot for useful discussions.
Thanks to Fabio Crameri and Dave May for constructive reviews.
%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
